---
title: "Regression Models Course Project"
author: "Daniel DeWaters"
date: "11/19/2019"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list=ls())
# Load dependencies/data
library(plyr, warn.conflicts = FALSE)
library(dplyr, warn.conflicts = FALSE)
library(ggplot2)
library(datasets)
data(mtcars)
```

## Executive Summary
This is an analysis for Motor Trend, a magazine about the automobile industry. Looking at a data set of a collection of cars, they are interested in exploring the relationship between a set of variables and miles per gallon (MPG) (outcome). They are particularly interested in the following two questions:

1. Is an automatic or manual transmission better for MPG?
2. Quantify the MPG difference between automatic and manual transmissions

## Cleaning
```{r}
# Convert am and vs columns into factors with meaningful labels
mtcars <-
  mtcars %>%
  mutate(vs = as.factor(vs)) %>%
  mutate(am = as.factor(am)) %>%
  mutate(cyl= as.factor(cyl))

levels(mtcars$vs) <- c("vshaped", "straight")
levels(mtcars$am) <- c("automatic", "manual")
```

## Exploratory Analysis
First we'll get a basic overview of all of the data
```{r exploratoryAnalysis1}
str(mtcars)
```

Then we'll get an overview of gas milage from automatic and manual transmission
```{r automaticSummary}
autoMPG <- 
  mtcars %>%
  filter(am == "automatic") %>%
  select(mpg)

manuMPG <-
  mtcars %>%
  filter(am == "manual") %>%
  select(mpg)

summary(autoMPG)
summary(manuMPG)

# Make boxplot
mpgBP <- ggplot(mtcars, aes(x=as.character(am), y=mpg, color=am))
mpgBP + geom_boxplot(aes(group=am)) +
  ggtitle("Boxplots of Gas Mileage from Automatic and Manual Transmissions") +
  xlab("Transmission") + ylab("Gas Milage (MPG") +
  theme(legend.position="none")
```

## Models
### Model 1
MPG is the outcome and am is the predictor
```{r model1}
fit1 <- lm(mpg~am, mtcars)
summary(fit1)
```
### Model 2
MPG is the outcome and all of the other variables are the predictors
```{r model2}
fit2 <- lm(mpg~., mtcars)
summary(fit2)
```

## Criteria
1. Did the student interpret the coefficients correctly?
3. Did the student fit multiple models and detail their strategy for model selection?
4. Did the student answer the questions of interest or detail why the question(s) is (are) not answerable?
5. Did the student do a residual plot and some diagnostics?
6. Did the student quantify the uncertainty in their conclusions and/or perform an inference correctly?
7. Was the report brief (about 2 pages long) for the main body of the report and no longer than 5 with supporting appendix of figures?